Building socket-aware BPF programs

### 2018-10-29 paper/paper.tex
paper: Flesh out introduction

\usepackage{xcolor}
\newcommand\todo[1]{\textcolor{red}{#1}}
\newcommand\linuxversion{v4.20}

@@

BPF, Linux, packet processing, sockets

@@

\begin{itemize}
\item BPF started with sockets, but only provided access to packets
\item Previous helper extensions allowed e.g. socket operation handling
\item Arbitrary socket lookup from socket/packet hooks allows more powerful BPF
      programs
\item Use cases revolve around building logic in socket/packet layer hooks that
      change the behaviour based on the existence and attributes of a socket.
\end{itemize}

@@

The implementation introduced in the 4.20 release cycle includes support for
looking up TCP and UDP sockets for IPv4 and IPv6 traffic, using the
\verb+SCHED_CLS+, \verb+SCHED_ACT+ and \verb+SK_SKB+ hook points. A proposed
patchset extends this to allow the same helpers to be used from the \verb+XDP+
hook points~\cite{sk-lookup-xdp}.

@@

@@

The core BPF maintainers provided great guidance and review of the patches, in
particular Alexei Starovoitov, Daniel Borkmann and Martin KaFai Lau. Additional
shout-out to Nitin Hande who provided testing and feedback.

>>>>

\usepackage{xcolor}
\newcommand\todo[1]{\textcolor{red}{#1}}
\newcommand\linuxversion{{v4.20}}

@@

BPF, firewalls, Linux, networks, packet processing, sockets

@@

The Berkeley Packet Filter~\cite{cbpf}, introduced in 1992, provided a model
for the implementation of flexible packet filtering by allowing runtime
modification of kernel logic from userspace applications. BPF quickly became
the de facto technology for monitoring the traffic flow within POSIX systems.
Originally, this was restricted to applying filtering upon sockets attached to
devices and not to implement filtering capabilities directly for traffic
flowing through a system. Subsequent work investigated using BPF as a basis for
packet filtering on devices~\cite{ipfw}, however this design does not appear to
have been integrated into major operating systems until the introduction of
extended BPF into Linux~\cite{lwnebpf}.

In recent years, multiple efforts have been made to build new filtering
implementations based upon
BPF~\cite{cilium,ovs-ebpf,nsx-harmony,droplet,xdp-ddos,alfwrapper}. Many of
these implementations aim to provide the functionality of a stateful
firewall~\cite{stateful-fw} - that is, to allow network policies to be defined
to allow traffic from one network endpoint to talk to another network endpoint,
and to implicitly allow responses to pass back through the firewall. Many
implementations over the past decade have developed this functionality in a
manner that is agnostic to the locality of the endpoints. This paper revisits
the possibilities for stateful firewalling when the endpoint socket is
co-located with the kernel providing firewalling functionality, effectively
combining the ideas of \cite{ipfw,stateful-fw,dfw}.

This paper is laid out as follows. Firstly, we discuss the constraints that the
BPF verifier places upon BPF programs and how those constraints translate into
requirements on the implementation of socket lookup helpers in the Linux BPF
API. The implementation of reference tracking functionality is presented.
Secondly, we describe extensions to the BPF API to provide socket introspection
capabilities, taking into account flexibility and performance. Thirdly, we
present some use cases for the new functionality, including more detail on the
stateful firewall case and a description of the use of the new functionality
for load-balancing. Finally, we conclude and assess the position of this work.

@@

The implementation introduced in the Linux~\linuxversion~release cycle includes
support for looking up TCP and UDP sockets for IPv4 and IPv6 traffic, using the
\verb+SCHED_CLS+, \verb+SCHED_ACT+ and \verb+SK_SKB+ hook points. A proposed
patchset extends this to allow the same helpers to be used from the \verb+XDP+
hook points~\cite{sk-lookup-xdp}.

@@

\todo{Early in the stack, figure out whether the traffic is destined for a
local process and if so, pass. Otherwise, perform load-balancing to another
node where the traffic is destined for.}

@@

The author would like to acknowledge Alexei Starovoitov, Daniel Borkmann,
Martin KaFai Lau, Nitin Hande and Thomas Graf for guidance, review and testing
of the patches.
